\section{Maria Aloni: Nothing is logical}

Formal semantics uses logic to analuze linguistic meaning.

Semantic data is explained by translation into logical language:

\[
    \text{natural language} \mapsto \text{logical language} \Rightarrow \text{models}
\]

Many issues:
\begin{enumerate}
    \item How to collect semantic data
    \item How to translate natural language expression
    \item Which logic to adopt
\end{enumerate}

Natural language inferences are genuinely semantic data.

In conversation we draw many inferences. Are all relevant for analysis of
linguistic meaning.

Grice's paradise:
\begin{enumerate}
    \item Sematnic innference: derivable by classical logic
    \item Pragmatic inference: derivable by conversational factors (as the
          product of rational interactions among cooperative language users)
\end{enumerate}

Goal of the project: provide formal account of third kind inferences which
captures quasi-semantic behaviour.

Strategy: decelop logic of conversation

Hypothesis: neglect-zero tendency is crucial pragmatic factor

\subsection{Challenge of free choice}

Deontic inference:
\[
    \text{You may go to the beach or the cinema}\to\text{You may go to the beach and you may go to the cinema}
\]

Epistemic inference:
\[
    \text{Mr X might be in victoria or in brixton}\to\text{mr x might be in victoria and he might be in brixton}
\]

Logical rendering:
\[
    \lozenge(p\vee q)\to(\lozenge a\wedge\lozenge b).
\]

Paradox of free choice: You may do (A or B) \(\to\) You may A

\[
    \lozenge (p\vee q)\to \lozenge p\>\>(\text{FC})
\]

But FC is not true because
\begin{enumerate}
    \item \(\lozenge p\)
    \item \(\lozenge p\vee q\)
    \item \(\lozenge q\)
\end{enumerate}
allows you to do anyting.

Natural language equivalent of the first step seems invalid: you may post this
letter \(\not\to\) you may post this letter or burn it.

Pragmatic solution: FC inferences are pragmatic inferences, step leading to 3 is
unjustified. We keep the logic as is.

Grammatical solution: FC inferences result from the application of covert
grammatical operator. We keep the logic as is.

Semantic solution: Change the logic to block the first step.

Present proposal: FC inferences as neglect zero effects derived for ``enriched''
formulas. This changes the logic.

FC and related inferences are:
\begin{enumerate}
    \item Not the result of conversational reasoning
    \item Not the effect of optional applications of grammatical application
    \item Straightforward consequence of neglect zero tendency:

          When interpreting a sentence, people neglect structures which verify
          the sentence by virtue of empty configuration.

          This tendency follows from the difficulty of a cognitive operation.

          Example: every square is black. When evaluating a sentence, we ignore
          models where it is vacuously true (models without squares).

          This explains why we see zero as a different natural number for
          example. Downward monotonic quantifiers are more costly to process than
          upward monotonic.
\end{enumerate}

\subsection{Classical vs team based modal logic}
Classical: formula true in one point

Team: formula true on subsection

\subsubsection{Bilateral state based modal logic}
Teams \(\to\) information states

Assertion and rejection conditions are modeled rather than truth:

\(M,s,\models \varphi\): \(\varphi\) is assertible in \(s\subseteq W\)

\(M,s,\dashv \varphi\): \(\varphi\) is rejectable in \(s\subseteq W\)

Bivalence fails in bsml on information states

\subsection{BSML}

\[
    \varphi := p \mid \neg\varphi \mid\varphi\vee\varphi\mid\varphi\wedge\varphi\mid\lozenge\varphi\mid \text{NE}.
\]

Models: classical Kripke models \((W,R,V)\).

States \(s\subseteq W\).
